A Switching Planner for Combined Task and Observation Planning

### 2010-11-29 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/intro.tex
Few new changes to the introduction.

Planning for agents acting in partially observable environments with
stochastic actions is extremely computationally
expensive \cite{mdp-complexity}. Even relatively small partially
observable Markov decision problems (POMDPs), represented by tens of
propositional variables, produce state spaces far larger than existing
optimal or close to optimal planners can handle. For the easier
problem of planning in completely observable stochastic domains, an
alternative approach has been to use classical sequential planners and
replan when actions have unplanned-for
outcomes \cite{yoon:etal:2007}. However, these approaches rely
crucially on being able to determine properties of the current state
in order to decide to replan. For partially observable domains, only
probabilistic information about these properties will in general be
available.

In this paper we describe, \pcogx, the planning component in a robotic
system that continuously deliberates in a stochastic dynamic
environment in order to achieve objectives set by the user, and
acquire knowledge about its surroundings. This domain has the
characteristics of a POMDP in that the state is not perfectly known,
and state information is gained by the robot through the use of
unreliable sensing actions. However, the state representation is
propositional, and even for relatively simple problems the underlying
model is too large for a POMDP solver to be applied.

Inspired by the replanning approaches used successfully for MDP
planning \cite{yoon:etal:2007,yoon:etal:2008} we propose an approach
that uses a replanning sequential planner to build the majority of the
plan, but uses a decision-theoretic planner on small subproblems where
reasoning about observations is required to find good plans. We refer
to the complete system as a {\em switching continual planner}. The
approach is not optimal, particularly as it is based on a satisficing
sequential planner, but it performs better than the sequential planner
alone and is fast enough to be used in real-time on a mobile robot for
decision-making.

As an illustration, consider a robot searching for a box that it knows
is often found in offices. The sequential planner might produce a plan
that assumes that {\sc Room a} is an office, that the box is in {\sc
Room a}, and that the box is in location 1 in {\sc Room a}. Given
these assumptions, the plan might be to go to {\sc Room a}, go to
location 1 and search for the box. When the plan gets executed and the
search begins the decision-theoretic planner is called since searching
involves making observations. The decision-theoretic planner might
retract the last assumption but leave the others, and build a plan to
search the whole of {\sc Room a} for the box. If this plan fails to
find the box, it will disprove the assumption that the box is in {\sc
Room a}, which will then lead to replanning by the sequential planner
based on the newly discovered knowledge.

To represent planning problems of this kind, we have developed a
first-order declarative language, called decision-theoretic planning
domain definition language (DTPDDL), for describing domains of
planning problems that correspond to POMDPs. We have created a system
that automatically constructs from this a deterministic planning domain
for the sequential planner that includes actions that allow
assumptions to be made about the values of imperfectly known state
variables. These are then used to restrict the state space and set the
reward function for the decision-theoretic planner, which can operate
with the original representation of the domain.

The sequential planner we use is based
on \fastdownward \cite{fast-downward}. We add the capability to
replan, and also allow agent knowledge to be represented explicitly,
so we can write actions that gain the agent knowledge. For the
decision-theoretic planner we use forward search in the information
state.

The switching planner approach gives us some important advantages in
our robotic domain. First, the main source of uncertainty in the
domain is unreliable observations, rather than stochasticity in the
action effects, and this approach is particularly suited to problems
that combine a deterministic task planning problem with observations
as the decision-theoretic planner is only invoked to determine the
values of variables. Secondly, the replanning makes the system very
robust to changing goals and the discovery of new facts about the
world, both of which are features of our domain.

>>>>

Planning for agents acting in partially observable environments with
stochastic actions is extremely computationally
expensive~\cite{mdp-complexity}. Even relatively small partially
observable Markov decision problems (POMDPs), represented by tens of
propositional variables, produce state spaces far larger than existing
optimal or close to optimal planners can handle. For the easier
problem of planning in completely observable stochastic domains, an
alternative approach has been to use classical sequential planners and
replan when actions have unplanned-for outcomes~\cite{yoon:etal:2007}.
Execution monitoring tracks properties of the state that determine
whether a new plan is necessary. For partially observable domains,
only probabilistic information about such properties will, in general,
be available.

%% However, these approaches rely
%% crucially on being able to determine properties of the current state
%% in order to decide to replan. For partially observable domains, only
%% probabilistic information about these properties will in general be
%% available.

This paper describes the planning component for a robotic system that
continuously deliberates in a stochastic dynamic environment in order
to achieve objectives set by the user, and acquire knowledge about its
surroundings. This domain features partial observability, particularly
because the state is not perfectly known, and state information is
gained by the robot through the use of unreliable sensing
actions. Using a propositionally factored state representation, for
interesting tasks the corresponding POMDP model is too large for a
POMDP solver to be applied directly. Inspired by the replanning
approaches used successfully for MDP
planning~\cite{yoon:etal:2007,yoon:etal:2008}, we propose an approach
that uses a replanning system to compute a reasonably valuable trace
in the model, and then uses a decision-theoretic planner on small
subproblems where reasoning about observations might be useful. We
refer to the complete system as a {\em switching continual
planner}. The approach is not optimal, particularly as it is based on
a satisficing sequential planner, but it performs better than the
sequential planner alone, and is fast enough to be used for real-time
decision-making on a mobile robot.

%% As an illustration, consider a robot searching for a box that it knows
%% is often found in offices. The sequential planner might produce a plan
%% that assumes that {\sc Room a} is an office, that the box is in {\sc
%% Room a}, and that the box is in location 1 in {\sc Room a}. Given
%% these assumptions, the plan might be to go to {\sc Room a}, go to
%% location 1 and search for the box. When the plan gets executed and the
%% search begins the decision-theoretic planner is called since searching
%% involves making observations. The decision-theoretic planner might
%% retract the last assumption but leave the others, and build a plan to
%% search the whole of {\sc Room a} for the box. If this plan fails to
%% find the box, it will disprove the assumption that the box is in {\sc
%% Room a}, which will then lead to replanning by the sequential planner
%% based on the newly discovered knowledge.

Our switching system is domain independent, taking domain and problem
descriptions in a first-order declarative language we have developed,
called the decision-theoretic planning domain definition language
(DTPDDL). From those descriptions of POMDPs, we automatically
construct a deterministic model for the sequential planner. That model
includes operators which correspond to assumptions about the values of
imperfectly known state variables. Assumptions scheduled by the
sequential system are used to propose a pragmatic abstract
belief-state space to the decision-theoretic system, and to modify the
reward function, so that system might pursue sensing related to those
assumptions.

%%  which can operate
%% with the original representation of the domain.

The sequential planner we use is based
on \fastdownward~\cite{fast-downward}. We add the capability to
replan, and also allow agent knowledge to be represented explicitly,
so we can write actions that gain the agent knowledge. LEASE CITE
MICHAEL's WORK AS THE BASIS FOR THIS. For the decision-theoretic
planner we have implemented our own forward search in the belief-state
space.

Our switching planner gives us some important advantages in our
robotic domain. First, the significant source of uncertainty in the
domain is the unreliability of observations, and this approach is
particularly suited to problems that combine a deterministic task
planning problem with observations as the decision-theoretic planner
is only invoked to determine a characteristic of the underlying
state. Secondly, the replanning makes the system very robust to
changing objectives and the discovery of new facts about the world,
both of which are features of our domain.

The remainder of the paper proceeds as follows. We begin by describing
POMDPs with a propositionally factored representation of states and
observations. We also outline the relationship between this and a flat
POMDP representation. Next we describe our DTPDDL language with
example declarations from a mobile robot exploration task. Then we
describe the switching planner, including how the sequential domain is
produced and a plan is generated, and how this then provides the input
for the decision-theoretic planner. Finally we provide an evaluation
of the system on the example robot domain, and finish with a short
survey of related work and some concluding remarks.
